---
title: "TidyTuesday 2"
author: "Cande Torres"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(tidytuesdayR)
library(ggplot2)
library(tidyverse)
library(devtools) 
library(openintro)
library(ggthemes)
library(gplots) 
library(RColorBrewer) 
```

```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 38)

billboard <- tuesdata$billboard
audio_features<- tuesdata$audio_features
```

```{r}
audio_2 <- audio_features %>% 
  mutate(ranking_type= 
           cut(spotify_track_popularity, 
               breaks = c(0,10,30,60,80,100),
               labels= c("High", "Above Average", "Intermetiate", "Bellow Average", "Low")))
```
```{r}
audio_2 %>% 
  count(spotify_track_explicit, ranking_type)
```

```{r}
audio_2 %>% 
  drop_na(spotify_track_explicit, ranking_type) %>% 
  ggplot() +
  geom_bar(aes(x = spotify_track_explicit, fill = ranking_type), position = "fill") +
  ylab("" ) +
  xlab("Explicit Track?") +
  scale_fill_viridis_d("Popularity") +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom")
```

After exploring the dataset variables I thought it would be really interesting to visualize the distribution of song popularity levels deciding whether they were explicit tracks or not. In spotify the songs in the top 10 are the most popular but the ranking runs until 100. By creating levels of popularity, one can see this distribution more clearly in the bar plots. This plot shows that in general, non-explicit songs are more popular on Spotify, which is something I didn't expect. 

Link to Data: https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/billboard.csv'
