---
title: "TidyTuesday1"
author: "Cande Torres"
date: "9/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

```{r message=TRUE}
library(maps) 
library(ggmap)
library(gplots) #new
library(RColorBrewer)
library(tidyverse)
library(tidytuesdayR)
library(dplyr)
library(jsonlite)
library(leaflet)
library(httr)
library(ggthemes)
library(sf)
```

```{r warning=FALSE}
driver_standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/driver_standings.csv')
```
```{r message=TRUE, warning=FALSE}
races <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/races.csv')
```

```{r message=TRUE}
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/drivers.csv')
```

```{r message=FALSE}
#download.file(
 # "http://ergast.com/downloads/f1db_ansi.sql.gz", 
  #destfile = "f1db-mysql.zip"
#)
```

```{r message=TRUE}
#Creat a more comprehensive data set
driver_results_df <- driver_standings %>% 
  left_join(races, by = "raceId") %>% 
  rename(driver_url = url) %>% 
  left_join(drivers, by = "driverId")
  
glimpse(driver_results_df)
```

```{r message=TRUE}
#Filter the data to only encompass the top 10 nationalities with more races won from 1950 to 2021.
driver_wins <- driver_results_df %>%
  group_by(nationality) %>% 
  summarise(total_wins = sum(wins)) %>% 
  arrange(desc(total_wins)) %>% 
  head(10)
```

```{r message=TRUE}
# generate labels for plot
lbl <- paste0(driver_wins$total_wins, "races")
driver_wins <- driver_wins %>% mutate(label = lbl)
```

```{r message=TRUE}
#Plot nationality by number of races won, add labels and color. 
driver_wins %>% 
  ggplot(aes(x=nationality,
             y=total_wins))+ 
  geom_col(fill="red")+
  theme_minimal() +
  labs(y="Number of Races won", x="Driver's nationality", title="Top-10 Nationalities of Race-Winners")+
  geom_text(aes(y = total_wins, label = label), vjust = -0.5)
```
This bar plot shows the top 10 nationalities of winners of Formula 1 races and the number of races won by each of them. I decided to plot it because in some countries races are really popular sports and a source of quality entertainment with many followers. Knowing the amount of races won by drivers of ones country appeals to national sentiment and pride. In the future, a good investigation would be to find the amount of supporters in each country to see if there is a correlation between number of national supporters and the number of wins achieved by drivers of that nationality. 

Link to Data: https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-09-07/readme.md

